---
title: "p8105_hw1_ah3108"
author: "Ava Hamilton"
date: "9/14/2019"
output: github_document
---


```{r}
library(tidyverse)
```

# Problem 1:

### Create dataframe
```{r create_df1}

hw1p1_df = tibble(
  rsamp = rnorm(8),
  log_vec = rsamp > 0,
  char_vec = c("a", "b", "c", "d", "e", "f", "g", "h"),
  fac_vec = factor(c("lev1", "lev3", "lev2", "lev2", "lev2", "lev1", "lev3", "lev3"))
  
)

```

### Take mean of variables in dataframe

```{r mean_Df}

mean(pull(hw1p1_df, rsamp))
mean(pull(hw1p1_df, log_vec))
mean(pull(hw1p1_df, char_vec))
mean(pull(hw1p1_df, fac_vec))

```

It was only possible to take the mean of the numeric variable "rsamp" (mean = `r round(mean(pull(hw1p1_df, rsamp)),3)`), and the logical variable "log_vec", (mean = `r round(mean(pull(hw1p1_df, log_vec)),3)`).


### Convert variables to numeric

```{r eval_df, eval = F}

as.numeric(pull(hw1p1_df, log_vec))
as.numeric(pull(hw1p1_df, char_vec))
as.numeric(pull(hw1p1_df, fac_vec))
```

a. The logical vector converts to 0s and 1s; able to take the mean of 0s and 1s in previous step.
b. The character vector will not convert to a numeric variable; cant take the mean in previous step.
c. The factor variable evaluates the input as a whole and assigns numeric values to each category alphabetically; even though categories can be numeric, taking the mean of a discrete categorical variable organized alphabetically is meaningless.



### Manipulate variables

```{r manip_vars}
as.numeric(pull(hw1p1_df, log_vec))*pull(hw1p1_df, rsamp)

as.factor(pull(hw1p1_df, log_vec))*pull(hw1p1_df, rsamp)

as.numeric(as.factor(pull(hw1p1_df, log_vec)))*pull(hw1p1_df, rsamp)
```



# Problem 2:

### Create dataframe

```{r create_df2}

hw1p2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  xyGT1_log = x + y > 1,
  num_log = as.numeric(xyGT1_log),
  fac_log = as.factor(xyGT1_log)
)
```

### Dataframe description

- The dataframe hw1p2_df has `r nrow(hw1p2_df)` rows and `r ncol(hw1p2_df)` columns. 
- The variable x is a random sample from a standard normal distribution of size 500. The variable x has a mean of `r mean(pull(hw1p2_df, x))`, a median of `r median(pull(hw1p2_df, x))`, and standard deviation of `r sd(pull(hw1p2_df, x))`.
- The proportion of cases for which `x + y > 1` is `r mean(pull(hw1p2_df, xyGT1_log))`.

### Scatterplots of y vs x

#### Logical color scale

```{r plotlog}

ggplot(hw1p2_df, aes(x = x, y = y, color = xyGT1_log)) + geom_point()


```

#### Numeric color scale
```{r plotnum}

ggplot(hw1p2_df, aes(x = x, y = y, color = num_log)) + geom_point()

```


#### Factor color scale
```{r plotfac}

ggplot(hw1p2_df, aes(x = x, y = y, color = fac_log)) + geom_point()

```

#### Comments

The color scales for the logical and factor variables are the same. Both have discrete colors representing the different levels. The color scale using the numeric vector shows a continuous color scale over the range of the numeric values, in this case 0 to 1. However, as there are only two values in the numeric vector, the plot is made up of two colors.


```{r export plot, include = F}

ggplot(hw1p2_df, aes(x = x, y = y, color = xyGT1_log)) + geom_point()


ggsave("firstScatter.pdf")

```


