---
title: "p8105_hw1_ah3108"
author: "Ava Hamilton"
date: "9/14/2019"
output: html_document
---


```{r}
library(tidyverse)
```

# Problem 1:

### Create dataframe
```{r}

hw1p1_df = tibble(
  rsamp = rnorm(8),
  log_vec = rsamp > 0,
  char_vec = c("a", "b", "c", "d", "e", "f", "g", "h"),
  fac_vec = factor(c("lev1", "lev3", "lev2", "lev2", "lev2", "lev1", "lev3", "lev3"))
  
)

```

### Take mean of variables in dataframe

```{r}

mean(pull(hw1p1_df, rsamp))
mean(pull(hw1p1_df, log_vec))
mean(pull(hw1p1_df, char_vec))
mean(pull(hw1p1_df, fac_vec))

```

It was only possible to take the mean of the numeric variable "rsamp" (mean = `r round(mean(pull(hw1p1_df, rsamp)),3)`), and the logical variable "log_vec", (mean = `r round(mean(pull(hw1p1_df, log_vec)),3)`).


### Convert variables to numeric

```{r, eval = F}

as.numeric(pull(hw1p1_df, log_vec))
as.numeric(pull(hw1p1_df, char_vec))
as.numeric(pull(hw1p1_df, fac_vec))
```

a. The logical vector converts to 0s and 1s; able to take the mean of 0s and 1s in previous step.
b. The character vector will not convert to a numeric variable; cant take the mean in previous step.
c. The factor variable evaluates the input as a whole and assigns numeric values to each category alphabetically; even though categories can be numeric, taking the mean of a discrete categorical variable organized alphabetically is meaningless.


